---
title: "Figures"
output: html_notebook
---

The packages used to create the figures

```{r,message=FALSE}
library(tidyverse)
library(colorspace)
#Figure 2: Current and wave directions from two current profilers
library(cowplot)
library(gridExtra)
#Figure 5: Proportion of time spent swimming in place against the current
library(lubridate)
library(circular)

#Figure 10: Distribution of welfare scores over time
library(readxl)
library(scales)

```

The data used to create the figures
```{r, message=FALSE}
#Figure 3: Current and wave directions from two current profilers; Figure 3: Wave height measured by two current profilers
sentstreym<-read_csv("sentinelv2.csv")
awacstreym<-read_csv("awaccurrent.csv")
awacalda<-read_csv("awacwaves.csv")
sentalda<-read_csv("sentwaves.csv")

#Figure 5: Depth measured by pressure sensors
deform<-read_csv("depths.csv")
colnames(deform)<- c("time","temp","Side west","Side north","Side south","Bottom west","Side east","Bottom east", "Bottom south","Bottom north")
deformlong<-pivot_longer(deform, cols= -c("time","temp"), names_to = "location", values_to = "depth")
sentsmall<-sentstreym[c(1,34,35,64,65)]
currentlong<-merge(deformlong,sentsmall, by.x = "time", by.y = "datetime")

currentlong<-currentlong %>% 
  mutate(direction=factor(ifelse(meandir<240 & meandir>90, "Southbound","Northbound"))) 

currentlong$location<-factor(currentlong$location, levels=c("Side west","Side south","Side east","Side north","Bottom west","Bottom south","Bottom east","Bottom north"))

#Figure 6: Proportion of time spent swimming in place against the current
southanalysis<-read_csv("nusentanalysis.csv")

sentshort<-sentsmall %>% 
  mutate(thirtymin=round_date(datetime, "30 minutes")) %>% 
  group_by(thirtymin) %>% 
  summarise(meanmag=mean(meanmag),meandir=mean(circular(meandir,units="degrees"))%%360)

behaviours2<-read_csv2("behaviours2.csv")
behaviours2$datetime<-as.POSIXct(paste(behaviours2$date,behaviours2$time), format="%d.%m.%Y %H:%M:%S")

behavioursent<-behaviours2 %>% 
  mutate(datetime=round_date(datetime,"5 minutes")) %>% 
  left_join(sentsmall, by=c("datetime"="datetime")) %>% 
  left_join(southanalysis, by=c("datetime"="datetime")) %>% 
  mutate(hourly=round_date(datetime, "hour")) %>% 
  left_join(awacalda, by=c("hourly"="hourly")) %>% 
  mutate(nonsome=none+some) %>% 
  mutate(somemany=some+many) %>% 
  mutate(manyshoal=many+shoal) %>% 
  mutate(any=somemany+shoal) %>% 
  mutate(total=none+some+many+shoal) %>%
  mutate(notmany=none+some+shoal) %>% 
  mutate(meanmag1=ifelse(is.na(meanmag),speedmean*1.5, meanmag)) %>% 
  mutate(meandir1=ifelse(is.na(meandir),dirmean, meandir)) %>% 
  mutate(direction=ifelse(meandir<250 & meandir>70,"Southbound","Northbound")) %>% 
  mutate(direction1=ifelse(meandir1<250 & meandir1>70,"Southbound","Northbound")) %>% 
  mutate(direction2=ifelse(dirmean<250 & dirmean>70,"Southbound","Northbound")) %>% 
  mutate(height=cut(Hm0,breaks=c(-Inf,1.4,Inf),labels=c("Small","Large"))) %>% 
  mutate(length=cut(Tp, breaks=c(-Inf,13,Inf),labels=c("Short","Long"))) %>% 
  mutate(streym=cut(meanmag1, breaks=c(-Inf,0.2,Inf),labels=c("Weak","Strong"))) %>% 
  mutate(biagainst=ifelse(against/(301-none)>0.5,1,0)) %>% 
  mutate(bimany=ifelse(manyshoal>150,1,0)) %>% 
  mutate(biany=ifelse(any>150,1,0)) %>% 
  mutate(wave=Hm0*Tp) %>% 
  mutate(noneprop=none/total) %>% 
  mutate(someprop=some/total) %>% 
  mutate(manyprop=many/total) %>% 
  mutate(shoalprop=shoal/total) %>% 
  mutate(nonesprop=nonsome/total) %>% 
  mutate(anyprop=any/total) 

levels(behavioursent$biagainst)<-c("Against current","Free swimming")


#Figure 7: Swimming effort measured in tail beats per second (tbps) over current
effort<-read_csv2("effort.csv")
effort$datetime<-as.POSIXct(effort$datetime, format = "%d.%m.%Y %H:%M")
effort$date<-as.Date(effort$datetime)
effort$camera<-factor(effort$camera)
effort$swimtype<-factor(effort$swimtype)
effort$against<-factor(effort$against)


effortsent<-effort %>% 
  mutate(thirtymin=floor_date(datetime, "30 minutes")) %>% 
  left_join(sentshort, by=c("thirtymin"="thirtymin")) %>% 
  left_join(southanalysis, by=c("datetime"="datetime"))%>% 
  mutate(hourly=floor_date(thirtymin, "hour")) %>% 
  left_join(awacalda, by=c("hourly"="hourly"))%>% 
  mutate(direction=ifelse(meandir<250 & meandir>70,"Southbound","Northbound")) %>% 
  mutate(direction2=ifelse(dirmean<250 & dirmean>70,"Southbound","Northbound")) %>% 
  mutate(height=cut(Hm0,breaks=c(-Inf,1.2,Inf),labels=c("Small","Large"))) %>% 
  mutate(length=cut(Tp, breaks=c(-Inf,12,Inf),labels=c("Short","Long"))) %>% 
  mutate(streym=cut(speedmean, breaks=c(-Inf,0.1,0.2,Inf),labels=c("Weak","Medium","Strong"))) %>% 
  mutate(tbps=3/effort)

levels(effortsent$against)<-c("Against current","Free swimming")

#Figure 9: 

collisiondat<-behavioursent %>% 
  mutate(netvis=ifelse(collision>0, "ja", netvis)) %>% 
  filter(netvis=="ja") %>% 
  filter(none < 300)%>% 
  mutate(nucols=collision/any) %>% 
  mutate(nucols2=collision/some)

collisiondat$nucols<-round(collisiondat$nucols,2)*300
collisiondat$nucols2<-round(collisiondat$nucols2,2)*300

#Figure 10: Depth and dispersal of fish over hour in the day when current is weak.

ekkoS<-read_csv("T1clean.csv")
ekkoN<-read_csv("T2clean.csv")

NEsouth<-ekkoN %>% 
  left_join(southanalysis, by=c("tenmin"="datetime")) %>% 
  left_join(sentsmall, by=c("tenmin"="fivemin")) %>% 
  mutate(direction=factor(ifelse(meandir<250 & meandir>70, "Southbound","Northbound"))) %>% 
  mutate(direction2=ifelse(dirmean<250 & dirmean>70,"Southbound","Northbound")) %>% 
   filter(hourly<"2020-02-15") %>% 
  mutate(streym=cut(meanmag, breaks=c(-Inf,0.12,0.22,Inf),labels=c("Weak","Medium","Strong"))) %>% 
  mutate(hourfac=factor(hour)) %>% 
  mutate(daynight=factor(ifelse(hour<15 & hour>8, "Day","Night")))

SEsouth<-ekkoS %>% 
  left_join(southanalysis, by=c("tenmin"="datetime")) %>% 
  left_join(sentsmall, by=c("tenmin"="fivemin")) %>% 
  mutate(direction=factor(ifelse(meandir<250 & meandir>70, "Southbound","Northbound"))) %>% 
  mutate(direction2=ifelse(dirmean<250 & dirmean>70,"Southbound","Northbound")) %>% 
   filter(hourly<"2020-02-15") %>% 
  mutate(streym=cut(meanmag, breaks=c(-Inf,0.12,0.22,Inf),labels=c("Weak","Medium","Strong"))) %>% 
  mutate(hourfac=factor(hour)) %>% 
  mutate(daynight=factor(ifelse(hour<15 & hour>8, "Day","Night")))


#Figure 11
ekkowsouth<-SEsouth %>% 
  left_join(awacalda, by=c("hourly"="hourly")) %>% 
  mutate(height=cut(Hm0,breaks=c(-Inf,1,Inf),labels=c("Small","Large"))) %>% 
  mutate(length=cut(Tp, breaks=c(-Inf,12,Inf),labels=c("Short","Long"))) 
ekkownorth<-NEsouth %>% 
  left_join(awacalda, by=c("hourly"="hourly")) %>% 
  mutate(height=cut(Hm0,breaks=c(-Inf,1,Inf),labels=c("Small","Large"))) %>% 
  mutate(length=cut(Tp, breaks=c(-Inf,12,Inf),labels=c("Short","Long"))) 

#Figure 12
welfare<-read_xlsx("Heilsukanningar.xlsx")
welfare<-welfare %>% 
  mutate(fins=as.ordered(fins)) %>% 
  mutate(eyepupil=as.ordered(eyepupil)) %>% 
  mutate(eyeiris=as.ordered(eyeiris)) %>% 
  mutate(snout=as.ordered(snout)) %>% 
  mutate(injury=as.ordered(injury)) %>% 
  mutate(cage=as.factor(cage)) %>% 
  mutate(injuries=ordered(injuries,levels=c("0","1","2","3","4","5","6","7","8","9"))) %>% 
  mutate(nucage=as.factor(nucage)) #Fish were moved from cage 5 to 6. This variable puts them in the same cage for easier direct comparison

```

The code used to create each figure

Figure 3: Current and wave directions from two current profilers

```{r,fig.width=9,fig.height=7}
source("ggplotwindrose2.R")
awaccurrentrose<-plot.windrose2(spd=awacstreym$meanmag, dir=awacstreym$meandir,spdmin=0,spdmax=0.6,spdres=.1)+theme(legend.position="none", axis.title.y = element_blank())+labs(title="AWAC Current")
sentvcurrentrose<-plot.windrose2(spd=sentstreym$meanmag, dir=sentstreym$meandir,spdmin=0,spdmax=0.6,spdres=0.1, legend="Current speed (m/s)")+labs(title="Sentinel V Current")+theme(axis.title.y = element_blank())
awacwaverose<-plot.windrose2(spd=awacalda$Hm0, dir=awacalda$DirTp,spdmin=0,spdmax=3.5,spdres=0.5)+theme(legend.position="none", axis.title.y = element_blank())+labs(title="AWAC Waves")
sentvwaverose<-plot.windrose2(spd=sentalda$Hs, dir=sentalda$Dp,spdmin=0,spdmax=3.5,spdres=0.5, legend="Wave height (m)")+labs(title="Sentinel V Waves")+theme(axis.title.y = element_blank(), plot.margin = margin(5,30,5,5,"pt"))
plot_grid(awaccurrentrose,sentvcurrentrose,awacwaverose,sentvwaverose,ncol=2, rel_heights = c(1,1), rel_widths= c(2,3))
#ggsave("windroses.pdf")
```
Figure 4: Wave height measured by two current profilers.

```{r,fig.width=7,fig.height=3}
sentwave<-ggplot(sentalda, aes(x=datetime, y=Hs, color="Hs"))+
  geom_point(size=0.5)+
  geom_point(aes(y=Hmax, color="Hmax"), size = 0.5)+
  labs(title="Sentinel V", x="Date",y="Wave height (m)", colour="Wave parameter")+
 scale_x_datetime(limits=as.POSIXct(c("2019-11-01","2020-03-01")))+
  lims(y=c(0,5))+
  guides(colour = guide_legend(override.aes =list(size=2)))+
  theme_classic()+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        legend.position=c(0.1,0.7))+
  scale_color_discrete_sequential(palette = "Magenta", l1=40, l2 = 80, c1=50,c2=50)
 

awacwave<-ggplot(awacalda, aes(x=datetime, y=Hm0, color = "Hs"))+
  geom_point(size = 0.5)+
  geom_point(aes(y=Hmax, color="Hmax"), size = 0.5)+
  labs(title="AWAC", x="Date",y="Wave height (m)")+
  scale_x_datetime(limits=as.POSIXct(c("2019-11-01","2020-03-01")),breaks="month")+
  theme_classic()+
   theme(legend.position = "none")+
  scale_color_discrete_sequential(palette = "Magenta", l1=40, l2 = 80, c1=50,c2=50)

plot_grid(sentwave,awacwave,ncol=1,rel_heights = c(1,1.3))

ggsave("waves.eps")
```

Figure 5: Depth measured by pressure sensors.

```{r,fig.width=7.5,fig.height=3}
ggplot(currentlong, aes(y = depth, x =meanmag, col=direction))+
  geom_point(alpha=1/70)+
  geom_smooth(method="lm", se=F)+
  facet_wrap(.~location, ncol=4)+
  labs(x= "Current speed (m/s)", y="Depth (m)", col="Current Direction")+
  ylim(-20,0)+
  scale_color_discrete_sequential(palette="Magenta", l1=40, l2 = 80, c1=50,c2=50)+
  theme_classic()+
  theme(panel.spacing.x = unit(4, "mm"))
ggsave("depths.tiff", dpi=300)
```

Figure 6: Proportion of time spent swimming in place against the current

```{r,fig.width=7,fig.height=3}
baplotfit2<-glm(biagainst~speedmean+direction2, data=subset(behavioursent, none<300), family=binomial)
baforplot2<-expand.grid(speedmean=seq(0,0.3,0.01),direction2=c("Northbound","Southbound"))
baforplot2$biagainst<-predict(baplotfit2, newdata=baforplot2, type="response")

ggplot(baforplot2, aes(x=speedmean, y=biagainst, colour=direction2))+
  geom_line(size=1)+
  geom_point(data=subset(behavioursent, none<300),aes(x=speedmean,y=against/any, col=direction2))+
  scale_color_discrete_sequential(palette="Magenta", l1=40, l2 = 80, c1=50,c2=50)+
  labs(x="Current speed (m/s)", y="Proportion swimming against current", color="Current direction")+
  theme_classic()
ggsave("against.eps")
```

Figure 7: swimming effort

```{r,fig.width=7,fig.height=3}
bigeffort<-lm(tbps~speedmean*length*direction2, data=subset(effortsent, against=="Against current"))
bigdata<-expand.grid(direction2=c("Southbound","Northbound"),speedmean=seq(0,0.3,0.01), length=c("Short","Long"))
predictedfree<-predict(bigeffort, newdata=bigdata, type="response", interval="confidence")
fit<-cbind(bigdata,predictedfree)

ggplot(fit, aes(x=speedmean,y=fit, col=length))+
  geom_line(size=1)+
  geom_ribbon(aes(ymin=lwr,ymax=upr,colour=NULL, fill=length),alpha=0.15)+
  geom_point(data=subset(effortsent, tbps<1.9), aes(x=speedmean, y=tbps, col=length))+
  labs(x="Current speed (m/s)",y="Swimming effort (tbps)", col="Wave length", fill="Wave length")+
  scale_fill_discrete_sequential(palette="Magenta",l1=40, l2 = 80, c1=50,c2=50)+
  scale_color_discrete_sequential(palette="Magenta",l1=40, l2 = 80, c1=50,c2=50)+
  ylim(0,2)+
  facet_grid(.~direction2)+
  theme_classic()
ggsave("effort.eps")
```

Figure 8: Proportion of time per video where "Many" (top panel) or "Any" (bottom panel) salmon were recorded
```{r,fig.width=7,fig.height=3}
manyboxplot<-ggplot(behavioursent, aes(x=camera, y=manyprop, color=height, fill=height))+
  geom_boxplot(outlier.shape=NA,alpha=0.4)+
  geom_point(position=position_jitterdodge())+
  scale_color_discrete_sequential(palette="Magenta", l1=40, l2 = 80, c1=50,c2=50)+
  scale_fill_discrete_sequential(palette="Magenta", l1=40, l2 = 80, c1=50,c2=50)+
  labs(x="Camera location",y="Proportion", col="Wave height", fill="Wave height")+
 # facet_wrap(.~streym)+
  theme_classic()
anyboxplot<-ggplot(behavioursent,aes(x=camera,y=anyprop,color=height, fill=height))+
  geom_boxplot(outlier.shape=NA,alpha=0.4)+
  geom_point(position=position_jitterdodge())+
  scale_color_discrete_sequential(palette="Magenta", l1=40, l2 = 80, c1=50,c2=50)+
  scale_fill_discrete_sequential(palette="Magenta", l1=40, l2 = 80, c1=50,c2=50)+
  labs(x="Camera location",y="Proportion", col="Wave height", fill="Wave height")+
#  facet_wrap(.~streym)+
  theme_classic()

grid.arrange(manyboxplot,anyboxplot)
horizontal<-arrangeGrob(manyboxplot,anyboxplot)
ggsave("horizontal.tiff",horizontal, dpi=300, height=3, width=7)
```

Figure 9: Number of net collisions per 300 seconds over wave height split by current speed

```{r,fig.width=7,fig.height=3}
ggplot(collisiondat, aes(x=Hm0,y=nucols, col=streym))+
  geom_point()+
  geom_smooth(method = "glm", 
        method.args = list(family = "poisson"), aes(fill=streym))+
  scale_color_discrete_sequential(palette="Magenta", l1=40, l2 = 80, c1=50,c2=50)+
  scale_fill_discrete_sequential(palette="Magenta", l1=40, l2 = 80, c1=50,c2=50)+
    labs(x="Wave height (m)", y="Collisions per 300 seconds",col="Current speed",fill="Current speed")+
  #facet_wrap(.~direction)+
  theme_classic()

ggsave("collisions.tiff", dpi=300)
```




Figure 10: depth and dispersal of fish over hour in the day when the current is weak

```{r, fig.width=7, fig.height=3}
SEsouth %>%
  subset(meanmag<0.12) %>%
  group_by(depth, hour,direction) %>% 
  summarise(meanSv=10*log10(mean(10^(Sv/10)))) %>% 
  ggplot(aes(x=hour,y=depth, fill=meanSv))+
  geom_tile()+
  labs(x="Hour", y="Depth (m)", fill="Mean Sv",title="South echo sounder")+
  scale_fill_gradientn(limits = c(-100,-70),
  colours=c("black", "darkmagenta", "lightpink", "white"))+
  theme_classic()
ggsave("hourdepth.eps")

```
Figure 11: depth and dispersal of fish over current

```{r, fig.width=7, fig.height=4}
allplots<-SEsouth %>%
  subset(daynight=="Day") %>%
  mutate(meanmag= round(meanmag, 2)) %>% 
  group_by(daynight,depth, meanmag,direction) %>% 
  summarise(meanSv=10*log10(mean(10^(Sv/10)))) %>% 
  drop_na() %>% 
  ggplot(aes(x=meanmag,y=depth, fill=meanSv))+
  geom_tile()+
  scale_fill_gradientn(limits = c(-105,-60),
  colours=c("black", "darkmagenta", "lightpink", "white"))+
  facet_grid(.~direction)+
  labs(title="South echosounder",x="", y="Depth (m)", fill="Mean Sv")+
  theme_classic()

allplotn<-NEsouth %>%
  subset(daynight=="Day") %>%
  mutate(meanmag= round(meanmag, 2)) %>% 
  group_by(daynight,depth, meanmag,direction) %>% 
  summarise(meanSv=10*log10(mean(10^(Sv/10)))) %>% 
  drop_na() %>%
  ggplot(aes(x=meanmag,y=depth, fill=meanSv))+
  geom_tile()+
  scale_fill_gradientn(limits = c(-105,-60),
  colours=c("black", "darkmagenta", "lightpink", "white"))+
  facet_grid(.~direction)+
  labs(title="North echosounder",x="Current speed (m/s)", y="Depth (m)", fill="Mean Sv")+
  theme_classic()

grid.arrange(allplots,allplotn)

ggsave("currentdepth.eps")
```

Figure 12: Depth and dispersal of fish over wave height (Hm0) in weak current
```{r, fig.width=7, fig.height=2}
wavesouth<-ekkowsouth %>%
  filter(daynight=="Night") %>% 
  filter(meanmag<0.12) %>%
  mutate(Hm0= round(Hm0, 1)) %>% 
  group_by(depth, Hm0) %>% 
   summarise(meanSv=10*log10(mean(10^(Sv/10)))) %>% 
   #subset(meanSv> -100) %>% 
  drop_na() %>%
  ggplot(aes(x=Hm0, y=depth, fill=meanSv)) +
  scale_fill_gradientn(limits = c(-100,-70),
  colours=c("black", "darkmagenta", "lightpink", "white"))+
  labs(title="South echosounder",x="Hm0 (m)", y="Depth (m)", fill="Mean Sv")+
  geom_tile()+
 # facet_wrap(.~direction)+
  theme_classic()

wavenorth<-ekkownorth %>%
  filter(daynight == "Night" ) %>%
  filter(meanmag<0.12) %>%
  mutate(Hm0= round(Hm0, 1)) %>% 
  group_by(depth, Hm0) %>% 
   summarise(meanSv=10*log10(mean(10^(Sv/10)))) %>% 
   #subset(meanSv> -100) %>% 
  drop_na() %>%
  ggplot(aes(x=Hm0, y=depth, fill=meanSv)) +
  scale_fill_gradientn(limits = c(-100,-70),
  colours=c("black", "darkmagenta", "lightpink", "white"))+
  labs(title="North echosounder",x="Hm0 (m)", y="Depth (m)", fill="Mean Sv")+
  geom_tile()+
#  facet_wrap(.~direction)+
  theme_classic()

grid.arrange(wavesouth,wavenorth,ncol=2)

ggsave("wavedepth.eps")
```

Figure 13: Distribution of welfare scores over time
```{r, fig.width=7, fig.height=3}
welfare %>% 
  filter(!(injuries %in% NA)) %>% 
   filter(date<"2020-06-15" ) %>% 
    count(date,nucage,injuries) %>% 
   ggplot(aes(x=date, y=n,fill = injuries))+
  geom_col(position ="fill")+#, width =1000000)+
  labs(y="Proportion", x="Date", fill="Total injuries")+
  scale_fill_discrete_sequential(palette="Magenta",drop=FALSE, l1=0, l2 = 100, c1=50,c2=50)+
  scale_x_datetime(breaks="2 months", labels = date_format("%m-%Y"))+
  theme_classic()+
  theme(axis.text.x=element_text(angle=45, vjust=1,hjust=1), panel.grid.major.y = element_line(color="grey80", linetype = 2))
ggsave("welfarescores.eps")
```

