---
title: "Statistical analysis"
output: html_notebook
---

Data analysis and models.

Packages
```{r}
library(lme4)
library(tidyverse)
library(circular)
library(lubridate)
library(lmerTest)
library(MASS)
```

Data

```{r}
#Used for most analyses
awacalda<-read_csv("awacwaves.csv")
southanalysis<-read_csv("nusentanalysis.csv")
sentstreym<-read_csv("sentvfivemins.csv")
sentsmall<-sentstreym[c(1,5,11,32,33)]

sentshort<-sentsmall %>% 
  mutate(thirtymin=round_date(fivemin, "30 minutes")) %>% 
  group_by(thirtymin) %>% 
  summarise(meanmag=mean(meanmag),meandir=mean(circular(meandir,units="degrees"))%%360, mag4=mean(mag4),dir4=mean(circular(dir4,units="degrees"))%%360)


# For measuring swimming effort
effort<-read_csv2("effort.csv")
effort$datetime<-as.POSIXct(effort$datetime, format = "%d.%m.%Y %H:%M")
effort$date<-as.Date(effort$datetime)
effort$camera<-factor(effort$camera)
effort$swimtype<-factor(effort$swimtype)
effort$against<-factor(effort$against)
effort<-effort[-7]

effortsent<-effort %>% 
  mutate(thirtymin=floor_date(datetime, "30 minutes")) %>% 
  left_join(sentshort, by=c("thirtymin"="thirtymin")) %>% 
  left_join(southanalysis, by=c("datetime"="datetime"))%>% 
  mutate(hourly=floor_date(thirtymin, "hour")) %>% 
  left_join(awacalda, by=c("hourly"="hourly"))%>% 
  mutate(direction=ifelse(meandir<250 & meandir>70,"Southbound","Northbound")) %>% 
  mutate(direction2=ifelse(dirmean<250 & dirmean>70,"Southbound","Northbound")) %>% 
  mutate(height=cut(Hm0,breaks=c(-Inf,1.2,Inf),labels=c("Small","Large"))) %>% 
  mutate(length=cut(Tp, breaks=c(-Inf,12,Inf),labels=c("Short","Long"))) %>% 
  mutate(streym=cut(speedmean, breaks=c(-Inf,0.1,0.2,Inf),labels=c("Weak","Medium","Strong"))) %>% 
  mutate(tbps=3/effort)

levels(effortsent$against)<-c("Against current","Free swimming")


# For other video observations
behaviours2<-read_csv2("behaviours2.csv")
behaviours2$datetime<-as.POSIXct(paste(behaviours2$date,behaviours2$time), format="%d.%m.%Y %H:%M:%S")
behaviours2<-behaviours2 %>% 
  mutate(thirtymin=round_date(datetime, "30 minutes"))

behavioursent<-behaviours2 %>% 
  left_join(sentshort, by=c("thirtymin"="thirtymin"))

behavioursent<-behavioursent %>% 
  mutate(datetime=round_date(datetime,"5 minutes")) %>% 
  left_join(southanalysis, by=c("datetime"="datetime"))

behavioursent<-behavioursent %>% 
  mutate(hourly=round_date(thirtymin, "hour")) %>% 
  left_join(awacalda, by=c("hourly"="hourly"))

behavioursent<-behavioursent %>% 
  mutate(nonsome=none+some) %>% 
  mutate(somemany=some+many) %>% 
  mutate(manyshoal=many+shoal) %>% 
  mutate(any=somemany+shoal) %>% 
  mutate(direction=ifelse(meandir<250 & meandir>70,"Southbound","Northbound")) %>% 
  mutate(direction2=ifelse(dirmean<250 & dirmean>70,"Southbound","Northbound")) %>% 
  mutate(height=cut(Hm0,breaks=c(-Inf,1.2,Inf),labels=c("Small","Large"))) %>% 
  mutate(length=cut(Tp, breaks=c(-Inf,12,Inf),labels=c("Short","Long"))) %>% 
  mutate(streym=cut(speedmean, breaks=c(-Inf,0.1,0.2,Inf),labels=c("Weak","Medium","Strong"))) %>% 
  mutate(biagainst=ifelse(against/(301-none)>0.5,1,0)) %>% 
  mutate(bimany=ifelse(manyshoal>150,1,0)) %>% 
  mutate(wave=Hm0*Tp)

#For collisions

collisiondat<-behavioursent %>% 
  mutate(netvis=ifelse(collision>0, "ja", netvis)) %>% #some net visibility wrongly coded as "no".
  filter(netvis=="ja") %>% 
  filter(none < 300) 
 
collisiondat<-collisiondat %>% 
 mutate(nucols=collision/any) #Collisions per seconds of fish visible

collisiondat$nucols<-round(collisiondat$nucols,2)*300 #extrapolated to 300 seconds of visible fish and rounded for purposes of poisson analysis

#For echo data

ekkoS<-read_csv("T1clean.csv")
ekkoN<-read_csv("T2clean.csv")

NEsouth<-ekkoN %>% 
  left_join(southanalysis, by=c("tenmin"="datetime")) %>% 
  left_join(sentsmall, by=c("tenmin"="fivemin")) %>% 
  mutate(direction=factor(ifelse(meandir<250 & meandir>70, "Southbound","Northbound"))) %>% 
  mutate(direction2=ifelse(dirmean<250 & dirmean>70,"Southbound","Northbound")) %>% 
   filter(hourly<"2020-02-15") %>% 
  mutate(streym=cut(meanmag, breaks=c(-Inf,0.12,0.22,Inf),labels=c("Weak","Medium","Strong"))) %>% 
  mutate(hourfac=factor(hour)) %>% 
  mutate(daynight=factor(ifelse(hour<15 & hour>8, "Day","Night"))) %>% 
  mutate(posv=Sv+120)

SEsouth<-ekkoS %>% 
  left_join(southanalysis, by=c("tenmin"="datetime")) %>% 
  left_join(sentsmall, by=c("tenmin"="fivemin")) %>% 
  mutate(direction=factor(ifelse(meandir<250 & meandir>70, "Southbound","Northbound"))) %>% 
  mutate(direction2=ifelse(dirmean<250 & dirmean>70,"Southbound","Northbound")) %>% 
   filter(hourly<"2020-02-15") %>% 
  mutate(streym=cut(meanmag, breaks=c(-Inf,0.12,0.22,Inf),labels=c("Weak","Medium","Strong"))) %>% 
  mutate(hourfac=factor(hour)) %>% 
  mutate(daynight=factor(ifelse(hour<15 & hour>8, "Day","Night"))) %>% 
  mutate(posv=Sv+120)

ekkowsouth<-SEsouth %>% 
  left_join(awacalda, by=c("hourly"="hourly")) %>% 
  mutate(height=cut(Hm0,breaks=c(-Inf,1,Inf),labels=c("Small","Large"))) %>% 
  mutate(length=cut(Tp, breaks=c(-Inf,12,Inf),labels=c("Short","Long"))) %>% 
  mutate(streym=cut(meanmag,breaks=c(-Inf,0.12,0.16,Inf),labels=c("Weak","Medium","Strong")))
ekkownorth<-NEsouth %>% 
  left_join(awacalda, by=c("hourly"="hourly")) %>% 
  mutate(height=cut(Hm0,breaks=c(-Inf,1,Inf),labels=c("Small","Large"))) %>% 
  mutate(length=cut(Tp, breaks=c(-Inf,12,Inf),labels=c("Short","Long"))) %>% 
  mutate(streym=cut(meanmag,breaks=c(-Inf,0.16,Inf),labels=c("Weak","Strong")))



```
Video observation data:

Swimming mode

```{r}
againstmod<-glmer(biagainst~direction2+speedmean+(1|camera),family=binomial,data=subset(behavioursent, none<300))
summary(againstmod)

```

Effort in relation to mode

```{r}
modemod<-lm(tbps~against, data=effortsent)
summary(modemod)
```

Effort in relation to waves and current

```{r}
effortmod<-lmer(tbps~speedmean:Tp+speedmean:direction2+speedmean+direction2+Tp+(1|against),data=effortsent)
summary(effortmod)
```


"Many" fish by camera

```{r}
cameramany<-lm(many~camera, data=behavioursent)
summary(cameramany)
```

"Many" fish by wave parameters

```{r}
manymodel<-lm(many~Hm0*Tp, data=behavioursent)
summary(manymodel)
anova(manymodel,test="F")
```

"Any" fish seen over camera and wave height

```{r}
anymodel<-lm(any~Hm0*camera, data=behavioursent)
summary(anymodel)
```

Collisions

```{r}
colmodel<-glm(nucols~Hm0*speedmean, faimly=poisson, data=collisiondat)
summary(colmodel)

```

Echo sounder analysis

Back scatter strength
```{r}
southmodfull<-lm(Sv~meanmag*direction,data=SEsouth)
summary(southmodfull)

northmodfull<-lm(Sv~meanmag*direction,data=NEsouth)
summary(northmodfull)

```

Residual analysis

```{r}
southresfull<-residuals(southmodfull)
southresmodfull<-lm(abs(southresfull)~SEsouth$meanmag*SEsouth$direction)
summary(southresmodfull)

northresfull<-residuals(northmodfull)
northresmodfull<-lm(abs(northresfull)~NEsouth$meanmag*NEsouth$direction)
summary(northresmodfull)

```

Depth analysis in relation to current

```{r}
southdepth<-lm(depth~meanmag*direction, weights=posv, data=subset(SEsouth, Sv>-90))
summary(southdepth)

northdpeth<-lm(depth~meanmag*direction, weights=posv, data=subset(NEsouth, Sv>-90))
summary(northdepth)

```

Depth in relation to waves

```{r}

southwave<-lm(depth~Hm0*direction+daynight, weights=posv, data=subset(ekkowsouth, Sv>-90))
summary(southwave)

northwave<-lm(depth~Hm0+direction+daynight, weights=posv, data=subset(ekkownorth, Sv>-90))
summary(northwave)
```

